%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% acknowledgements.tex
%% NOVA thesis document file
%%
%% Text with acknowledgements
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE acknowledgements.tex}%

\begin{ntacknowledgements}

I am deeply grateful to my adviser Professor João Leitão, who enabled and guided me throughout both the preparation and the elaboration of this dissertation, sharing his wisdom and guiding me along the process. Under his guidance, I was able to develop and validate innovative solutions and contribute to this scientific field.

I would also like to express my utmost gratitude to my co-advisor João Bordalo, with whom I met almost weekly to discuss results, problems and new ways of innovating during the development of this thesis. His guidance and experience in the subject truly enabled me to provide some of my best work.

I also wish to give a huge thanks to my friends! The ones from before University Diogo and Alexandre, who have since then been a part of my life. Also a huge thanks to the friends that I have met through this course, specially Pedro, Tomás, André and Guilherme who have been there almost since the start and always supported me during these five years and during the development of this dissertation. I also wish to give a huge thanks to my friends from the music world, especially João, José and Clara, who were responsible for granting me the opportunity to make music with them and make this process much less stressful.

I owe my heartfelt gratitude to my parents, who enabled me to have this opportunity and privilege of cultivating my potential and pursuing my studies. Furthermore, I also wish to express my gratitude to the rest of my family who have always been there along the way encouraging me, especially my grandmother Natividade, who always believed in me and gave me the strength I needed every single day. Finally, I wish to express my utmost appreciation for my partner Tatiana, who stood with me through thick and thin during the last years of this course, and during this dissertation. Especially, my parents, my grandmother and my partner shaped who I am today.

\end{ntacknowledgements}